% !TEX encoding = UTF-8 Unicode
%!TEX root = thesis.tex
% !TEX spellcheck = en-US
%%=========================================
\chapter[Summary]{Summary and Recommendations for Further Work}
Better prior MCMC (Improved prior)$=>$ higher effective sample size.\\
GEV instead of POT! (since AR-GARCH $\sim$ i.i.d).\\
Improved ACER program. ACER continious CI so qq-plot is possible (given a 1\% event, where on the extreme prediction confidence level is that)\\
Automatically selecting u for POT MCMC each turn.\\
REMEMBER IN THEORY, WHY ADAPTIVE MCMC!! MAX EFFECTIVE SAMPLE SIZE, AND MORE IMPORTANTLY AUTOMATIC!!!!! MINIMAL HUMA INTERACTION.\\

%%This is the last chapter
%In this final chapter you should sum up what you have done and which results you have got. You should also discuss your findings, and give recommendations for further work.

%%=========================================
\section{Summary and Conclusions}
Here, you present a brief summary of your work and list the main results you have got. You should give comments to each of the objectives in Chapter 1 and state whether or not you have met the objective. If you have not met the objective, you should explain why (e.g., data not available, too difficult).

This section is similar to the Summary and Conclusions in the beginning of your report, but more detailed---referring to the the various sections in the report.

%%=========================================
\section{Discussion}
Here, you may discuss your findings, their strengths and limitations.
%%=========================================
\section{Recommendations for Further Work}
You should give recommendations to possible extensions to your work. The recommendations should be as specific as possible, preferably with an objective and an indication of a possible approach.

The recommendations may be classified as:
\begin{itemize}
\item Short-term
\item Medium-term
\item Long-term
\end{itemize}